Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-06-26 src/tex/ms.tex
update

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Distribution of training and validation NR waveforms in the
	$q-\chi_{PN}$ space. The blue stars represent the waveforms used to compute
	the training loss and the orange dots represent the waveforms used for
	validation. \kw{It is hard to see the blue stars on top of the orange. Play
	around with the size of the dots and change the colors. Also add the extra
	NR waveforms in table 2 with a different color but use the stars.}}}
	\label{fig:intrin_space}
\end{figure}

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio is defined as
	$q=m_1/m_2\geq 1$ and the spins are denoted by $\chi_{1,2}$. Out of the 11 waveforms listed
	here, nine of them are also used in the original IMRPhenomD calibration. The remaining two
	are SXS analogues to two of the BAM NR simulations used in the original calibration
	\citep{khan2016frequency}. \kw{BE EXPLICIT ABOUT WHAT DO YOU MEAN THE REMAINING TWO!}}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional NR waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

@@

We divided the parameter space into four regions to analyze the effect of the
recalibration procedure on each region separately
(Fig.~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
this scenario are listed in Tab.~\ref{tab:quadrants} and the loss functions 
are calculated using a simple average of the mismatches, $\mathcal{L}_{\mathrm{ave}}$. The top-left and
bottom-right regions have limited data for $q>4$, hence the results are only
valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
equal-spin waveforms lying on the diagonal improve the most. 
\kl{The improvement above the diagonal is also significant comparing with the original waveform model. However, the improvement are not as profound as optimizing all waveform at once. This shows that the ansatz can achieve a better fit using training waveforms with similar spins, even for waveforms with non-equal spin parameters. }
%The improvement above the diagonal is also significantly more than when optimizing using
%all waveforms at once. This can be seen clearly in the top panel of Fig.~\ref{fig:all_quadrants}, where we see
%that the optimized histogram shifts to lower mismatches.
\footnote{
	Note that the top panel of Fig.~\ref{fig:all_quadrants} specifically shows the
	distribution of mismatches for the top-left quadrant of Fig.~\ref{fig:ps_q148_quadrant}.
} On the other hand, the
improvement below the diagonal is negligible, indicating that our method struggles to
further improve over the original IMRPhenomD waveform. 
Interestingly, we can see in the bottom-right quadrant that the waveform does not show a significant change
to the overall mismatch (see bottom panel of Fig.~\ref{fig:all_quadrants}). 
This suggests that the IMRPhenomD waveform ansatz fits waveforms with
$\chi_1>0$ and $\chi_2<0$ less well.

>>>>

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Distribution of training and validation NR waveforms in the
	$q-\chi_{PN}$ space. The blue stars represent the waveforms used to compute
	the training loss and the orange dots represent the waveforms used for
	validation.}
	\label{fig:intrin_space}
\end{figure}

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio is defined as
	$q=m_1/m_2\geq 1$ and the spins are denoted by $\chi_{1,2}$. Out of the 11 waveforms listed
	here, two SXS waveforms are analogues to two of the BAM NR simulations used in the original calibration (SXS:BBH:1417 as A7 and SXS:BBH:1418 as A9 in \citep{khan2016frequency}). The rest of the SXS waveforms are also used in the original IMRPhenomD calibration.}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule Code         & $q$ & $\chi_1$ & $\chi_2$ \\
		\midrule\midrule SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional NR waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

@@

We divided the parameter space into four regions to analyze the effect of the
recalibration procedure on each region separately
(Fig.~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
this scenario are listed in Tab.~\ref{tab:quadrants} and the loss functions 
are calculated using a simple average of the mismatches, $\mathcal{L}_{\mathrm{ave}}$. The top-left and
bottom-right regions have limited data for $q>4$, hence the result is only valid up to $q\leq4$ and we take waveforms with $q\leq4$ as validation waveforms in these two regions. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
equal-spin waveforms lying on the diagonal improve the most. 
\kl{The improvement above the diagonal is also significant comparing with the original waveform model and it has a similar improvement as optimizing all waveform at once. Notice that optimizing over all waveforms gives a slightly better improvement. This shows that the ansatz can achieve a better fit using training waveforms with similar spins, even for waveforms with non-equal spin parameters.}
%The improvement above the diagonal is also significantly more than when optimizing using
%all waveforms at once. This can be seen clearly in the top panel of Fig.~\ref{fig:all_quadrants}, where we see
%that the optimized histogram shifts to lower mismatches.
\footnote{
	Note that the top panel of Fig.~\ref{fig:all_quadrants} specifically shows the
	distribution of mismatches for the top-left quadrant of Fig.~\ref{fig:ps_q148_quadrant}.
} On the other hand, the improvement below the diagonal is negligible, indicating that our method struggles to
further improve over the original IMRPhenomD waveform. 
Interestingly, we can see in the bottom-right region that the waveform does not show a significant change
to the overall mismatch (see bottom panel of Fig.~\ref{fig:all_quadrants}). 
This suggests that the IMRPhenomD waveform ansatz fits waveforms with
$\chi_1>0$ and $\chi_2<0$ less well.
